---
title: "Crude oil trading model"
output: html_notebook
---

```{r}

library(tidyverse)
library(Rblpapi)
library(xgboost)
library(zoo)
library(data.table)
library(TTR)
library(lubridate)

```


```{r}

blpConnect()

cl1 <- bdh(securities = "CL1 COMB Comdty", fields = c("PX_LOW", "PX_HIGH",  "PX_LAST"), start.date = today()-years(100)) %>% 
  
cl1 %>% filter(is.na(PX_LOW)|is.na(PX_HIGH))

```

#feature engineering
```{r}
library(TTR)
library(zoo)

macd <- 
TTR::MACD(cl1$PX_LAST, percent = T) %>% 
  as.tibble() %>% 
  mutate(date=cl1$date) %>% 
  select(date, signal)

rsi <- 
TTR::RSI(cl1$PX_LAST) %>% 
  as.tibble() %>%
  mutate(date=cl1$date) %>% 
  rename(rsi=1)  

uo_dat <- cl1 %>% filter(date>'2000-01-01') %>% select(date, PX_HIGH, PX_LOW, PX_LAST)
ultimate_oscillator <- 
  TTR::ultimateOscillator(HLC = uo_dat[-1], n = c(7, 14, 28), wts = c(4, 2, 1)) %>% 
  as.tibble() %>% 
  mutate(date = uo_dat$date)
rm(uo_dat)

ultimate_oscillator%>%ggplot()+geom_line(aes(date, value))

realized_vol <- cl1 %>%
  select(date, PX_LAST) %>% 
  mutate(vol30 = PX_LAST/lag(PX_LAST),
         vol30 =  rollapplyr(vol30, FUN = sd, width = 30, fill = NA),
         vol30 = vol30*sqrt(260),
         vol90 = PX_LAST/lag(PX_LAST),
         vol90 =  rollapplyr(vol90, FUN = sd, width = 90, fill = NA),
         vol90 = vol90*sqrt(260)) %>% 
  mutate(vol_variance = (vol90*100)/(vol30*100)-1)  %>% 
  select(-PX_LAST)

onemonth <- 
bdh(securities = "CL1 Comdty", fields = c("1ST_MTH_IMPVOL_90.0%MNY_DF", "1ST_MTH_IMPVOL_110.0%MNY_DF"), start.date = today()-years(100)) %>%  mutate(onemodif = `1ST_MTH_IMPVOL_110.0%MNY_DF`/`1ST_MTH_IMPVOL_90.0%MNY_DF`-1) 
  
twomonth <- 
bdh(securities = "CL1 Comdty", fields = c("2ND_MTH_IMPVOL_90.0%MNY_DF", "2ND_MTH_IMPVOL_110.0%MNY_DF"), start.date = today()-years(100)) %>%  mutate(twomodif = `2ND_MTH_IMPVOL_110.0%MNY_DF`/`2ND_MTH_IMPVOL_90.0%MNY_DF`-1) 

threemonth <- 
bdh(securities = "CL1 Comdty", fields = c("3MTH_IMPVOL_90.0%MNY_DF", "3MTH_IMPVOL_110.0%MNY_DF"), start.date = today()-years(100)) %>%  mutate(threemodif = `3MTH_IMPVOL_110.0%MNY_DF`/`3MTH_IMPVOL_90.0%MNY_DF`-1) #%>%
  # ggplot() + 
  # geom_line(aes(date, difference))


ma_matrix <- cl1 %>% select(1, 4) %>% 
  mutate(levelchange5 = PX_LAST-lag(PX_LAST,5),
         levelchange15 = PX_LAST-lag(PX_LAST,15),
         levelchange50 = PX_LAST-lag(PX_LAST,50),
         levelchange100 = PX_LAST-lag(PX_LAST,100),
         levelchange200 = PX_LAST-lag(PX_LAST, 200),
         ma5 = SMA(PX_LAST, 5), 
         ma15 = SMA(PX_LAST, 15), 
         ma50 = SMA(PX_LAST, 50), 
         ma100 = SMA(PX_LAST, 100), 
         ma200 = SMA(PX_LAST, 200)) %>% 
  select(-PX_LAST)

technical_indicators <- 
  macd %>% 
  full_join(rsi, by = "date") %>% 
  full_join(realized_vol, by = "date") %>%
  melt(id.var = 1) %>%
  group_by(variable) %>% 
  mutate(change = value/lag(value, 10)) %>% 
  rename(variable1=variable) %>%
  melt(id.var = c(1, 2)) %>% 
  mutate(variable = paste(variable1, variable, sep = "_")) %>% 
  dcast(date~variable, value.var = "value") %>% 
  full_join(ma_matrix, by = "date")
  # full_join(onemonth) %>%
  # full_join(twomonth) %>%
  # full_join(threemonth)
  
technical_indicators %>% 
  melt(id.var=1) %>% 
  ggplot() +
  geom_line(aes(date, value)) +
  facet_wrap(variable~., scales = "free")


```



#Macro variables
```{r}

weekly_crude_inventories <- 
bdh(securities = 'DOEASCRD Index', fields = c("ECO_RELEASE_DT", "PX_LAST"), start.date = today()-years(100)) %>%
  mutate(date=ECO_RELEASE_DT) %>%
  na.omit() %>% 
  select(-ECO_RELEASE_DT) %>% 
  rename(crude_inventory=2) %>% 
  mutate(release = paste("Release", 1:nrow(.)))

blpConnect()

'CONCCONF Index' %>%
  bdh(securities = ., fields = c("ECO_RELEASE_DT", "PX_LAST"), start.date = today()-years(100)) %>% 
  mutate(date=ECO_RELEASE_DT) %>%
  na.omit() %>% 
  select(-ECO_RELEASE_DT) %>% 
  rename(crude_inventory=2) %>% 
  mutate(release = paste("Release", 1:nrow(.)))


weekly_crude_inventories %>% 
  ggplot() +
  geom_line(aes(date, crude_inventory))

bds(security = "DOESCRUD Index",  field = "ECO_RELEASE_DT_LIST")

```

#Tanker data
```{r}

alt_tickers <- 
tribble(
~name, ~ticker,  
'vessels_in_service', 'VESLCOIS Index',
'dwt_crude_vessels', 'VESLCODW Index',
'vessels_on_order', 'VESLCOOO Index', 
'vessel_orderbook', 'VESLCOPC Index'
)

alt_data <- NULL  

for(i in 1:nrow(alt_tickers)){
  alt_data <- 
    bdh(securities = alt_tickers$ticker[i], fields = c("PX_LAST"), start.date = today()-years(100)) %>% 
    mutate(name = alt_tickers$name[i]) %>%
    rbind(alt_data)
  }

alt_data %>% 
  group_by(name) %>% 
  # mutate(change = PX_LAST/lag(PX_LAST)-1) %>% 
  ggplot() +
  geom_line(aes(date, PX_LAST)) +
  facet_wrap(name~., scales = "free")


```



# model 
```{r}

dataset <- 
cl1 %>% select(date, PX_LAST) %>% rename(y=PX_LAST) %>% 
  mutate(week = paste("Week ",week(date), sep = ""),
         month = month(date, label = T, abbr = F)) %>% 
  mutate(key = 1) %>%
  spread(key = month, value = key, fill = 0) %>% 
  mutate(key = 1) %>%
  spread(key = week, value = key, fill = 0) %>% 
  full_join(technical_indicators) %>% 
  na.omit() %>% 
  # tail() %>% 
  select(-y)

dataset%>%tail()

```

# Model - build a long / short signal. When tomorrow's close is >0, long, when <0, short 
```{r}

dataset 

signal <- 
cl1 %>% select(date, PX_LAST) %>% rename(y=PX_LAST) %>% 
  mutate(price_change = y/lag(y)-1) %>% 
  na.omit() %>% 
  mutate(y = lead(y),
         price_change = lead(price_change),
         tradesignal = case_when(
           price_change==0~0,
           price_change<0~ -1,
           price_change>0~ 1
         )) %>% 
  select(date, tradesignal=y)

model_data <- 
dataset %>%
  inner_join(signal) %>%
  na.omit()

tomorrow <- 
dataset %>%
  inner_join(signal) %>%
  # na.omit() %>% 
  filter(is.na(tradesignal)) %>%
  select(-tradesignal)

tomorrow_xgb <- xgb.DMatrix(data =  as.matrix(tomorrow[,-1]))

subset <-  sample(seq_len(nrow(model_data)), size = floor(0.75 * nrow(model_data)))
train <- model_data[subset,]
y_fit <- train["tradesignal"]
train <- train%>%select(-tradesignal)
train_xgb <- xgb.DMatrix(data =  as.matrix(train[,-1]), label = y_fit$tradesignal)

test <- model_data[-subset,]
y_test <- test["tradesignal"]
test <- test%>%select(-tradesignal)
test_xgb <- xgb.DMatrix(data =  as.matrix(test[,-1]), label = y_test$tradesignal)

cv.nround <- 1000
cv.nfold  <- 5

mdcv <- xgb.cv(data=train_xgb, nthread=6, nfold=cv.nfold, nrounds=cv.nround, verbose = T)

min_logloss <-  mdcv$evaluation_log %>% filter(test_rmse_mean==min(test_rmse_mean))

xgboostmodel <-  xgboost(data = train_xgb, nrounds = min_logloss$iter)

xgboost::xgb.importance(model = xgboostmodel)

predict(object = xgboostmodel, newdata = test_xgb) %>% 
  as.tibble() %>% 
  mutate(date = test$date) %>% 
  inner_join(cl1%>%select(1,4)) %>%
  mutate(resid = value/PX_LAST-1) %>%
  ggplot() +
  geom_line(aes(date, value), color = "red") +
  geom_line(aes(date, PX_LAST), color = "blue") 

predict(xgboostmodel, newdata = tomorrow_xgb)



```


```{r}

fitted_values <- NULL

for(i in 252:nrow(model_data)){

train <- model_data %>% dplyr::slice(1:i)
y_fit <- train["tradesignal"]
train <- train%>%select(-tradesignal)
train_xgb <- xgb.DMatrix(data =  as.matrix(train[,-1]), label = y_fit$tradesignal)

tmw <- i+1

if(tmw>nrow(model_data)){
  break
}

tomorrow <- model_data %>% dplyr::slice(tmw) %>% select(-tradesignal)
tomorrow_xgb <- xgb.DMatrix(data =  as.matrix(tomorrow[,-1]))

cv.nround <- 1000
cv.nfold  <- 5

mdcv <- xgb.cv(data=train_xgb, nfold=cv.nfold, nrounds=cv.nround, verbose = T, early_stopping_rounds = 50)
min_logloss <-  mdcv$evaluation_log %>% filter(test_rmse_mean==min(test_rmse_mean))
xgboostmodel <-  xgboost(data = train_xgb, nrounds = min_logloss$iter)

fitted_values <- 
predict(xgboostmodel, newdata = tomorrow_xgb) %>% 
  as.tibble() %>% 
  mutate(date = tomorrow$date) %>% 
  rbind(fitted_values)

}


fitted_values

```





















